### 2022-09-10 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@d199f6d27a498bbadc447c73c51cd36074336e9e 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Solving the \(n\)-language problem: A
ecologist's guide to learning Julia}
\vskip 5em

%
\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{3,2}

\textsuperscript{1}\,McGill University\quad \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences\quad \textsuperscript{3}\,Université de
Montréal

\textbf{Correspondance to:}\\
Michael D. Catchen --- \texttt{michael.catchen@mcgill.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill
Julia is a good language, ecologists should learn it.

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

\hypertarget{outline}{%
\section{Outline}\label{outline}}

An ecologists guide to learning Julia

\begin{itemize}
\tightlist
\item
  Why should ecologists learn Julia?

  \begin{itemize}
  \tightlist
  \item
    No point ignoring the elephant in the room: R is by far the most
    used language in ecology
  \item
    But many of the problems in contemporary ecology require
    functionality that R just can't handle (mass individual-based
    models, big data management, quick numerical methods)
  \item
    \texttt{julia} has particular features that solve a broad chunk of
    the difficulties of computational analysis of ecological data

    \begin{itemize}
    \tightlist
    \item
      and improves user experience by enabling code that is
      \emph{idiomatic}.
    \end{itemize}
  \item
    this makes julia well suited to become "{[}a platform for ecological
    prediction and forecasting{]}; McIntyre etal, Urban etal
  \end{itemize}
\item
  In what order should people approach new topics

  \begin{itemize}
  \tightlist
  \item
    understand the core concepts that make julia different from other
    languages first

    \begin{itemize}
    \tightlist
    \item
      Types
    \item
      multiple dispatch
    \end{itemize}
  \item
    Learn data management in Julia (emphasize analogies to tidyverse)

    \begin{itemize}
    \tightlist
    \item
      DataFrames
    \end{itemize}
  \item
    Learn stats/ml in Julia (emphasize anologies to GLM in R and
    scikit-learn/tensorflow in python)
  \item
    Learn visualization (Makie)
  \end{itemize}
\item
  Discussion

  \begin{itemize}
  \tightlist
  \item
    Talk about adjacent scientific computing course
  \item
    What does the future of computing in ecology look like?
  \end{itemize}
\end{itemize}

\begin{longtable}[]{@{}l@{}}
\toprule
\endhead
\begin{minipage}[t]{0.05\columnwidth}\raggedright
- Why should ecologists learn julia? - Well, there are the criteria that
are directly measurable that make it better than R/Python: - fast -
native support on GPUs - But there are also criteria that are more
subjective, and that take experience and practice using the language to
appreciate - clever use of dispatch patterns - use of one-lienrs - using
parameterized types well - You will learn how to be a better programmer
in \emph{any} language, because smart use of julia requires
understanding some fundemantal concepts in programming that are `hidden'
from users in R/python - The biggest reason \emph{not} to use julia is
that the ecology/evolution package ecosystem in R is larger, and the ML
ecosystem in python is more popular. However: - you can call \emph{any}
R/python function/library using RCall/PyCall in julia - More packages
isn't necessarily better when they don't work together\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

In order to measure, understand, and mitigate the consequences of
anthropogenic change on ecosystems the services they provide, ecologists
need a modern set of computational tools (Urban \emph{et al.} 2022).

These tools must be performant, but crucially modular and interfaceable
(McIntire \emph{et al.} 2022)

The so-called ``two-language problem'' in computational science, where
it is easier for a researcher to developer a prototype of a
model/simulation in a high-level language, like Python or R, and later
have to port the model to a lower-level compiled language because the
performance of these compiled languages (e.g.~C++/Fortran) is orders of
magnitude faster than high level interpreted languages. In fact, many of
the most popular tools in higher-level languages are actually thin
wrappers around a compiled (often C++) base (e.g. tidyverse, keras,
numpy, TensorFlow, scikit-learn, pandas, etc.). However, the skills
required to use or debug---let alone write---scientific software in
these lower level langauges is not often taught.

Ecological data is often difficult to access and reuse (Gonzalez \&
Peres-Neto 2015; Poisot \emph{et al.} 2019). Synthesizing data into a
single product suitable for analysis often remains tedious as data are
not in formats that can be easily combined or interfaced.

Here we propose that we can solve this problem through standardization
(Zimmerman 2008)---developing a common definition such that data
collected in a variety of contexts can be assimilated while minimizing
the overhead of data cleaning and wrangling. A common representation of
ecological data will have three primary benefits: it will \textbf{1})
enable new forms of analysis by making it easier to combine data from
different sources (Heberling \emph{et al.} 2021), \textbf{2)} enable
continuous integration of new data for next-generation biodiversity
monitoring (Kühl \emph{et al.} 2020), and \textbf{3)} aid in open
sharing and reproducability of published results (Zimmerman 2008;
Borregaard \& Hart 2016). Here, we briefly review approaches to data
standardization developed in other fields, in order to determine what
makes an open standard succeed in promoting data sharing, and what
doesn't. Based on the properties of good standards we identify, we
propose building a living standard for ecological data in the
\texttt{Julia} programming language, and argue this is necessary to
obtain the three primary benefits of standardization mentioned earlier.

We propose that that Julia has certain properties absent in other
popular languages for scientific computing that make it particularly
suited for the development of a cohesive, modular, and extendible set of
tools ideal for the development of a platform for ecological analysis
{[}@{]}.

\hypertarget{the-nature-of-computation-in-julia}{%
\section{The nature of computation in
Julia}\label{the-nature-of-computation-in-julia}}

\begin{figure}
\centering
\includegraphics{./figures/multiple_dispatch.png}
\caption{TODO: caption. Adapted from Karpinski 2019 ``The unreasonable
effectiveness of multiple dispatch''}
\end{figure}

\hypertarget{types}{%
\subsection{Types}\label{types}}

Why is this useful for ecologists? Often times in ecology, the same
information is represented in different formats. Two packages in R might
not agree on what the ``correct'' format to represent information is.

At the core of the Julia language is its \emph{type system}. Type
systems can often be alienating to those who learned programming in
so-called \emph{dynamically} typed languages (like R, python, and
JavaScript). In dynamically-typed languages, \texttt{x=\ 5}, and
\texttt{x\ =\ "hello\ world"} and the language won't care that you
changed the type of information that was stored in \texttt{x} from a
number to a string. Practically, this form of dynamic-typing was adopted
because it is far more convenient to write code like that above than
defining variables with explicit types, e.g.~how you would in C:
\texttt{char\ c\ =\ "a";} and \texttt{int\ x\ =\ 5;}.

Julia doesn't require explicit type declarations, meaning
\texttt{x\ =\ 5} is perfectly valid code, but internallyJulia is doing
the bookkeeping of what type of information is stored in \texttt{x},
from an \texttt{Int64} to a \texttt{String} in the above example.

Using explicit types is central to Julia's speed, but also enables much
of its most unique and user-friendly functionality, primarily the use of
a \emph{multiple-dispatch} system.

\hypertarget{dispatch}{%
\subsection{Dispatch}\label{dispatch}}

\emph{Dispatch} refers to the way a computer program decides what
function to call.

In many staticly-typed languages, you are allow to use the same function
name more than once.

\hypertarget{doing-computational-science-in-julia}{%
\section{\texorpdfstring{Doing computational science in
\texttt{Julia}}{Doing computational science in Julia}}\label{doing-computational-science-in-julia}}

\hypertarget{managing-data}{%
\subsection{Managing Data}\label{managing-data}}

DataFrames.jl and DFMeta.

\hypertarget{doing-statistics-and-machine-learning}{%
\subsection{Doing statistics and machine
learning}\label{doing-statistics-and-machine-learning}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  Learn about the statistics ecosystem: StatsBase, Statistics, GLM, MLJ,
  Flux, Turing
\end{enumerate}

\hypertarget{doing-simulation}{%
\subsection{Doing simulation}\label{doing-simulation}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{7}
\tightlist
\item
  Learn about the simulation libraries (DiffEq, DynamicGrids)
\item
  Learn how various statistics/simulation libraries work togethe
\end{enumerate}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Defining a living standard for ecological data in \texttt{Julia} will
make it easier to combine data from different sources by splitting the
process of data aggregation from the process of analysis. Integrating
data from a particular study, or a new database, would be as simple as
implementing the interface from the data source to the standardized
types. Data from individual studies could be incorporated into public
repositories containing both the raw data and the interface to Julia
data structures, and this combined data/interface package is all that is
needed to either reproduce the results or incorporate that particular
study's data into analysis. This will make combining data from multiple
sources easier, and yield benefits for the development and
implementation of novel methods, as the software for analysis becomes
separate from the software for data cleaning and aggregation.

We envision a modern set of tools for ecology in \texttt{Julia} based
around the standardized types. Far outside of ecology, the term
``ecosystem'' is used metaphorically to describe a set of software tools
that work together. We imagine multiple ``trophic-levels'' of packages
for ecological science in \texttt{Julia} based around the ``basal'' set
of standardized types --- a modular set of tools that can be chained
together create arbitrarily complex analysis pipelines. that can be
scaled to meet the needs of next-generation biodiversity monitoring.

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Borregaard2016MorRep}{}%
Borregaard, M.K. \& Hart, E.M. (2016). Towards a more reproducible
ecology. \emph{Ecography}, 39, 349--353.

\leavevmode\hypertarget{ref-Gonzalez2015ActSta}{}%
Gonzalez, A. \& Peres-Neto, P.R. (2015). Act to staunch loss of research
data. \emph{Nature}, 520, 436--436.

\leavevmode\hypertarget{ref-Heberling2021DatInt}{}%
Heberling, J.M., Miller, J.T., Noesgaard, D., Weingart, S.B. \& Schigel,
D. (2021). Data integration enables global biodiversity synthesis.
\emph{Proceedings of the National Academy of Sciences}, 118.

\leavevmode\hypertarget{ref-Kuhl2020EffBio}{}%
Kühl, H.S., Bowler, D.E., Bösch, L., Bruelheide, H., Dauber, J.,
Eichenberg, David., \emph{et al.} (2020). Effective Biodiversity
Monitoring Needs a Culture of Integration. \emph{One Earth}, 3,
462--474.

\leavevmode\hypertarget{ref-McIntire2022PerRei}{}%
McIntire, E.J.B., Chubaty, A.M., Cumming, S.G., Andison, D., Barros, C.,
Boisvenue, C., \emph{et al.} (2022). PERFICT: A Re-imagined foundation
for predictive ecology. \emph{Ecology Letters}, 25, 1345--1351.

\leavevmode\hypertarget{ref-Poisot2019EcoDat}{}%
Poisot, T., Bruneau, A., Gonzalez, A., Gravel, D. \& Peres-Neto, P.
(2019). Ecological Data Should Not Be So Hard to Find and Reuse.
\emph{Trends in Ecology \& Evolution}, 34, 494--496.

\leavevmode\hypertarget{ref-Urban2022CodLif}{}%
Urban, M.C., Travis, J.M.J., Zurell, D., Thompson, P.L., Synes, N.W.,
Scarpa, A., \emph{et al.} (2022). Coding for Life: Designing a Platform
for Projecting and Protecting Global Biodiversity. \emph{BioScience},
72, 91--104.

\leavevmode\hypertarget{ref-Zimmerman2008NewKno}{}%
Zimmerman, A.S. (2008). New Knowledge from Old Data: The Role of
Standards in the Sharing and Reuse of Ecological Data. \emph{Science,
Technology, \& Human Values}, 33, 631--652.

\end{CSLReferences}

\end{document}